---
title: "FARS functions"
author: "Anne Mickan"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Overview of Fars functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, echo = FALSE,include=FALSE}
library(dplyr)
library(readr)
library(tidyr)
library(maps)
library(graphics)
library(Fars)
```

#### `Fars` TUTORIAL 
### How to use the `Fars` functions. 

The package includes functions for reading, summarizing and graphing data from the current working directory. 
The data are from the Fatality Analysis Reporting System, and have to be structured in this specific way for the 
functions to work. ([FARS](https://www.nhtsa.gov/research-data/fatality-analysis-reporting-system-fars)).

## *Create a file name*

All files read in and processed by functions in this package should follow the following naming convention:
         
 accident_*year*.csv.bz2

To create file names based on a given *year*, use `make_filename(year)`.

```{r make_filename}
# creating file name for the year 2013
name_2013 <- make_filename(2013)
name_2013
```
## *Import data*

Once a filename has been created, you can read the data for that year in by using `fars_read`. 
```{r fars_read}
# creating file name for year 2014
name_2014 <- make_filename(2014)
setwd(system.file("data", package = "Fars"))
# reading in data from 2014
data_2014 <- fars_read(name_2014)
# show the first 6 rows of the data
head(data_2014)
```

## *Sumarize data*

For one or more years, you can get a summary display of the number of observations per month for each year with the function `fars_summarize_years(years)`

```{r fars_summarize_years}
years <- c(2013, 2014, 2015)
setwd(system.file("data", package = "Fars"))
sum_years <- fars_summarize_years(years)
sum_years
```


## *Graph data*

The `fars_map_state()` function creates a map of a chosen state and plots the number of casualties reported within a chosen year *year* for that *state* geographically onto the state map. 

```{r fars_map_state, fig.align='center'}
setwd(system.file("data", package = "Fars"))
fars_map_state(4, 2013)
fars_map_state(10, 2015)
```
